---
title: "Geospatial data in R"
subtitle: "Workshop in Tallinn university on Jan 25th 2023"
author: Markus Kainu
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-location: right
    theme: simplex
---

## Pre-workshop assignment

Thank you for signing up for Wednesday afternoon data workshop.

To complete the hands on demo at the end, you need a computer and for that we have two options: [posit.cloud](https://posit.cloud) -platform or your own with late version of [R](https://www.r-project.org/)-software (R \> 4.0) installed.

If you are new to R, I recommend registering to [posit.cloud](https://posit.cloud) as it will just work. If you are more experienced with R or you want to have things as your things, please go on and install [R](https://www.r-project.org/), [git](https://git-scm.com/) and an IDE for R. AS IDE we will use Posits [RStudio](https://posit.co/download/rstudio-desktop/), but you can just as well use [vscode](https://code.visualstudio.com/docs/languages/r) or [Emacs](https://ess.r-project.org/).

### Registering to posit.cloud

You can register to [posit.cloud](https://posit.cloud) either by creating an account using email or using your existing Github or Google accounts. After signing up, log in to the system.

Workshop material is hosted at Gitlab <https://gitlab.com/muuankarski/tallinn_geodata_20230125/> and at post.cloud you should create a new project from it. 

### Pre-workshop assignment

Fire up your preferred R environment. Create a **new project** from **git repository** and paste this Gitlab repository as a source `https://gitlab.com/muuankarski/tallinn_geodata_20230125/`.

Finally, paste these two *code chunks* to your R console

*Chunk 1: Install packages and check if it worked*

```{r, message = FALSE, warning = FALSE, eval = FALSE}
pkgs <- c("sf","ggplot2","geofi","dplyr","tidyr","mapview","leaflet","httr2","fs")
install.packages(pkgs)
installed_pkgs <- rownames(installed.packages())
if (all(pkgs %in% installed_pkgs)) print("Success") else print("Try installing again or change to posit.cloud!")
```

*Chunk 2: Submit your session info to the instructor using a web api*

```{r, message = FALSE, warning = FALSE, eval = FALSE}
library(httr2)
req <- request("http://91.145.79.96:3310/workshop_check")
sessi <- sessionInfo()
lista <- list(Rver = sessi$R.version,
              platform = sessi$platform,
              locale = sessi$locale,
              running = sessi$running,
              installed_sf= "sf" %in% installed_pkgs,
              installed_geofi= "geofi" %in% installed_pkgs,
              installed_geofi= "leaflet" %in% installed_pkgs
              )
req <- req %>% req_body_json(lista)
resp <- req_perform(req)
resp_body_string(resp)
```

**The process above, logging into Posit.cloud, creating a project from a git repository and running the pre-workshop assignments, is shown in the short video below!**

<video controls src="video/tallinn_pre.mp4" width = "100%"></video>

## Workshop on Wednesday

### Slides

<iframe class="slide-deck" src="slides/" width="100%" height="660px">

</iframe>

### Hands-on demo

We will interactively go through the script [`workshop.R`](https://gitlab.com/muuankarski/tallinn_geodata_20230125/-/blob/master/workshop.R) using you preferred computing environment.
